#Reproducible Research: Peer Assessment 1
========================================================

This document contains code and outcomes for the first Peer Assessment assignment for the course Reproducible Research. The data used is from an activity monitoring device.

## Loading and preprocessing the data
```{r Loading}
fullData <- read.csv("activity.csv", stringsAsFactors=F)
install.packages("date")
library(date)
fullData$date <- as.Date(fullData$date, format = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r meanSteps }
stepsByDate <- by(fullData$steps, fullData$date, sum)
hist(stepsByDate, xlab="Steps", main="Total Steps by Day", xlim=c(0,25000), breaks=8)
 
myMean <- mean(stepsByDate, na.rm=TRUE)
myMedian <- median(stepsByDate, na.rm=TRUE)

```
The mean total number of steps per day is `r myMean`
The median number of steps per day is `r myMedian`

## What is the average daily activity pattern?
``` {r dailyActivity}
stepsByInterval <- aggregate(steps~interval, data=fullData, mean, na.rm=TRUE)
plot(stepsByInterval$steps~stepsByInterval$interval, type="l", ylab="Daily Average Steps", xlab="Interval")
mostSteps <- stepsByInterval$interval[which.max(stepsByInterval$steps)]
```
The interval with the maximum number of steps is `r mostSteps`

## Imputing missing values
```{r imputeMissing}
totalNA <- sum(is.na(fullData$steps)
              
imputedDF <- fullData
mySteps <- mean(stepsByInterval$steps)
imputedDF$steps[is.na(imputedDF$steps)] <- mySteps

imputedStepsByDate <- by(imputedDF$steps, imputedDF$date, sum)
hist(imputedStepsByDate, xlab="Steps", main="Total Steps by Day (Imputed)", xlim=c(0,25000), breaks=8)
 
imputedMean <- mean(imputedStepsByDate, na.rm=TRUE)
imputedMedian <- median(imputedStepsByDate, na.rm=TRUE)

```
The mean total number of steps per day is `r imputedMean`
The median number of steps per day is `r imputedMedian`
The values of the original and imputed data do not differ much. There is not much impact.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdayWeekend}
fullData$dayName = weekdays(fullData$date)
fullData$dayType <- if(fullData$dayName %in% c("Saturday","Sunday")) {"weekend"} else {"weekday"}
```
Unfortunately, I do not have time to finish this last question because the assignment is due and I need to get it pushed to github.
